\chapter{Filtering}

This chapter introduces the most commonly used filters found in the toolkit.
Most of these filters are intended to process images. They will accept one or
more images as input and will produce one or more images as output. ITK is
based on a data pipeline architecture in which the output of one filter is
passed as input to another filter. (See Section
\ref{sec:DataProcessingPipeline} on page \pageref{sec:DataProcessingPipeline}
for more information.)


\section{Thresholding}
\ifitkFullVersion
\label{sec:ThresholdingFiltering}
\fi

The thresholding operation is used to change or identify pixel values based
on specifying one or more values (called the \emph{threshold} value). The
following sections describe how to perform thresholding operations using
ITK.

\subsection{Binary Thresholding}
\label{sec:BinaryThresholdingImageFilter}

\ifitkFullVersion
\input{BinaryThresholdImageFilter.tex}
\fi

\subsection{General Thresholding}
\label{sec:ThresholdingImageFilter}

\ifitkFullVersion
\input{ThresholdImageFilter.tex}
\fi


\section{Edge Detection}

\subsection{Canny Edge Detection}
\ifitkFullVersion
\input{CannyEdgeDetectionImageFilter.tex}
\fi


\section{Casting and Intensity Mapping}
\label{sec:CastingImageFilters}

The filters discussed in this section perform pixel-wise intensity mappings.
Casting is used to convert one pixel type to another, while intensity mappings
also take into account the different intensity ranges of the pixel types.

\subsection{Linear Mappings}
\label{sec:IntensityLinearMapping}

\ifitkFullVersion
\input{CastingImageFilters.tex}
\fi

\subsection{Non Linear Mappings}
\label{sec:IntensityNonLinearMapping}

The following filter can be seen as a variant of the casting filters. Its main
difference is the use of a smooth and continuous transition function of
non-linear form.

\ifitkFullVersion
\input{SigmoidImageFilter.tex}
\fi
  

\section{Gradients}
\label{sec:GradientFiltering}

Computation of gradients is a fairly common operation in image processing. The
term ``gradient'' may refer in some contexts to the gradient vectors and in
others to the magnitude of the gradient vectors. ITK filters attempt to
reduce this ambiguity by including the \emph{magnitude} term when
appropriate. ITK provides filters for computing both the image of gradient
vectors and the image of magnitudes.

\subsection{Gradient Magnitude}
\label{sec:GradientMagnitudeImageFilter}

\ifitkFullVersion
\input{GradientMagnitudeImageFilter.tex}
\fi

\subsection{Gradient Magnitude With Smoothing}
\label{sec:GradientMagnitudeRecursiveGaussianImageFilter}

\ifitkFullVersion
\input{GradientMagnitudeRecursiveGaussianImageFilter.tex}
\fi


\subsection{Derivative Without Smoothing}
\label{sec:DerivativeImageFilter}

\ifitkFullVersion
\input{DerivativeImageFilter.tex}
\fi


\section{Neighborhood Filters}
\label{sec:NeighborhoodFilters}

The concept of locality is frequently encountered in image processing in the
form of filters that compute every output pixel using information from a small
region in the neighborhood of the input pixel.  The classical form of
these filters are the $3 \times 3$ filters in 2D images. Convolution masks
based on these neighborhoods can perform diverse tasks ranging from noise
reduction, to differential operations, to mathematical morphology.

The Insight toolkit implements an elegant approach to neighborhood-based image
filtering.  The input image is processed using a special iterator called the
\doxygen{NeighborhoodIterator}. This iterator is capable of moving over all the
pixels in an image and, for each position, it can address the pixels in a local
neighborhood. Operators are defined that apply an algorithmic operation in the
neighborhood of the input pixel to produce a value for the output pixel.  The
following section describes some of the more commonly used filters that take
advantage of this construction. (See Chapter
\ref{sec:ImageIteratorsChapter} on page
\pageref{sec:ImageIteratorsChapter} for more information about iterators.)

\subsection{Mean Filter}
\label{sec:MeanFilter}

\ifitkFullVersion
\input{MeanImageFilter.tex}
\fi

\subsection{Median Filter}
\label{sec:MedianFilter}

\ifitkFullVersion
\input{MedianImageFilter.tex}
\fi


\subsection{Mathematical Morphology}
\label{sec:MathematicalMorphology}

Mathematical morphology has proved to be a powerful resource for image
processing and analysis \cite{Serra1982}. ITK implements mathematical
morphology filters using NeighborhoodIterators and
\doxygen{NeighborhoodOperator}s.  The toolkit contains two types of image
morphology algorithms, filters that operate on binary images and filters that
operate on grayscale images.

\subsubsection{Binary Filters}
\label{sec:MathematicalMorphologyBinaryFilters}

\ifitkFullVersion
\input{MathematicalMorphologyBinaryFilters.tex}
\fi


\subsubsection{Grayscale Filters}
\label{sec:MathematicalMorphologyGrayscaleFilters}

\ifitkFullVersion
\input{MathematicalMorphologyGrayscaleFilters.tex}
\fi




\section{Smoothing Filters}
\label{sec:SmoothingFilters}

Real image data has a level of uncertainty that is manifested in the
variability of measures assigned to pixels. This uncertainty is usually
interpreted as noise and considered an undesirable component of the image
data. This section describes several methods that can be applied to reduce
noise on images.

\subsection{Blurring}
\label{sec:BlurringFilters}

Blurring is the traditional approach for removing noise from images. It is
usually implemented in the form of a convolution with a kernel. The effect of
blurring on the image spectrum is to attenuate high spatial
frequencies.  Different kernels attenuate frequencies in different ways. One
of the most commonly used kernels is the Gaussian. Two implementations of
Gaussian smoothing are available in the toolkit. The first one is based on a
traditional convolution while the other is based on the application of IIR
filters that approximate the convolution with a Gaussian
\cite{Deriche1990,Deriche1993}.

\subsubsection{Discrete Gaussian}
\label{sec:DiscreteGaussianImageFilter}

\ifitkFullVersion
\input{DiscreteGaussianImageFilter.tex}
\fi


\subsubsection{Binomial Blurring}
\label{sec:BinomialBlurImageFilter}

\ifitkFullVersion
\input{BinomialBlurImageFilter.tex}
\fi

\subsubsection{Recursive Gaussian IIR}
\label{sec:RecursiveGaussianImageFilter}

\ifitkFullVersion
\input{SmoothingRecursiveGaussianImageFilter.tex}
\fi


\subsection{Local Blurring}
\label{sec:BlurringFunctions}

In some cases it is desirable to compute smoothing in restricted regions of the
image, or to do it using different parameters that are computed locally.  The
following sections describe options for applying local smoothing in images.

\subsubsection{Gaussian Blur Image Function}
\label{sec:GaussianBlurImageFunction}

\ifitkFullVersion
\input{GaussianBlurImageFunction.tex}
\fi

\subsection{Edge Preserving Smoothing}
\label{sec:EdgePreservingSmoothingFilters}

\subsubsection{Introduction to Anisotropic Diffusion}
\label{sec:IntroductionAnisotropicDiffusion}
\ifitkFullVersion
\input{AnisotropicDiffusionFiltering.tex}
\fi


\subsubsection{Gradient Anisotropic Diffusion}
\label{sec:GradientAnisotropicDiffusionImageFilter}

\ifitkFullVersion
\input{GradientAnisotropicDiffusionImageFilter.tex}
\fi



\subsubsection{Curvature Anisotropic Diffusion}
\label{sec:CurvatureAnisotropicDiffusionImageFilter}

\ifitkFullVersion
\input{CurvatureAnisotropicDiffusionImageFilter.tex}
\fi

\subsubsection{Curvature Flow}
\label{sec:CurvatureFlowImageFilter}

\ifitkFullVersion
\input{CurvatureFlowImageFilter.tex}
\fi

\subsubsection{MinMaxCurvature Flow}
\label{sec:MinMaxCurvatureFlowImageFilter}

\ifitkFullVersion
\input{MinMaxCurvatureFlowImageFilter.tex}
\fi


\subsubsection{Bilateral Filter}
\label{sec:BilateralImageFilter}

\ifitkFullVersion
\input{BilateralImageFilter.tex}
\fi



\subsection{Edge Preserving Smoothing in Vector Images}
\label{sec:VectorAnisotropicDiffusion}

Anisotropic diffusion can also be applied to images whose pixels are vectors.
In this case the diffusion is computed independently for each vector
component.  The following classes implement versions of anisotropic diffusion
on vector images.

\subsubsection{Vector Gradient Anisotropic Diffusion}
\label{sec:VectorGradientAnisotropicDiffusionImageFilter}

\ifitkFullVersion
\input{VectorGradientAnisotropicDiffusionImageFilter.tex}
\fi

\subsubsection{Vector Curvature Anisotropic Diffusion}
\label{sec:VectorCurvatureAnisotropicDiffusionImageFilter}

\ifitkFullVersion
\input{VectorCurvatureAnisotropicDiffusionImageFilter.tex}
\fi



\subsection{Edge Preserving Smoothing in Color Images}
\label{sec:ColorAnisotropicDiffusion}

\subsubsection{Gradient Anisotropic Diffusion}
\label{sec:ColorGradientAnisotropicDiffusion}

\ifitkFullVersion
\input{RGBGradientAnisotropicDiffusionImageFilter.tex}
\fi

\subsubsection{Curvature Anisotropic Diffusion}
\label{sec:ColorCurvatureAnisotropicDiffusion}

\ifitkFullVersion
\input{RGBCurvatureAnisotropicDiffusionImageFilter.tex}
\fi



\section{Distance Map}
\label{sec:DistanceMap}

\ifitkFullVersion
\input{DanielssonDistanceMapImageFilter.tex}
\fi



\section{Geometric Transformations}
\label{sec:GeometricalTransformationFilters}

\subsection{Filters You Should be Affraid to Use}

\label{sec:ScaryImageFilters}
\subsection{Change Information Image Filter}

This one is the scariest and more dangerous filter in the entire toolkit. You
should not use this filter unless you are entirely certain that you know what
you are doing. In fact if you decide to use this filter, you should write your
code, then go for a long walk, get more coffee and ask yourself if you really
needed to use this filter. If the answer is yes, then you should discuss this
issue with someone you trust and get his/her opinion in writing.  In general,
if you need to use this filter, it means that you have a poor image provider
that is putting your career at risk along with the life of any potential
patiens whose images you may endup processing.

\subsection{Flip Image Filter}

\ifitkFullVersion
\input{FlipImageFilter.tex}
\fi

\subsection{Resample Image Filter}
\label{sec:ResampleImageFilter}

\subsubsection{Introduction}

\ifitkFullVersion
\input{ResampleImageFilter.tex}
\fi

\subsubsection{Importance of Spacing and Origin}
\ifitkFullVersion
\input{ResampleImageFilter2.tex}
\fi

\subsubsection{A Complete Example}
\ifitkFullVersion
\input{ResampleImageFilter3.tex}
\fi

\subsubsection{Rotating an Image}
\ifitkFullVersion
\input{ResampleImageFilter4.tex}
\fi

\subsubsection{Rotating and Scaling an Image}
\ifitkFullVersion
\input{ResampleImageFilter5.tex}
\fi

\subsubsection{Typical uses of resampling}

These section describes two very common uses of resampling.

First subsampling a volume in order to reduce its memory size, second resample
the volume in order to have isotropic pixel spacings.

In both cases, it is required to apply a low-pass filter before resampling the
image. The purpose of this filter is to attenuate the content of high spatial
frequencies on the input image and therefore prevent aliasing artifacts during
the resampling. 

\ifitkFullVersion
\input{SubsampleVolume.tex}
\fi


\ifitkFullVersion
\input{ResampleVolumesToBeIsotropic.tex}
\fi



