
\chapter{Segmentation}

Segmentation of medical images is a challenging task. A myriad of different
methods have been proposed and implemented in recent years. In spite of the
huge effort invested in this problem, there is no single approach that can
generally solve the problem of segmentation for the large variety of image
modalities existing today.

The most effective segmentation algorithms are obtained by carefully
customizing combinations of components. The parameters of these components are
tuned for the characteristics of the image modality used as input and the
features of the anatomical structure to be segmented.

The Insight Toolkit provides a basic set of algorithms that can be used to
develop and customize a full segmentation application. Some of the most
commonly used segmentation components are described in the following
sections.


\section{Region Growing}

Region growing algorithms have proven to be an effective approach for image
segmentation. The basic approach of a region growing algorithm is to start
from a seed region (typically one or more pixels) that are considered to be
inside the object to be segmented. The pixels neighboring this region are
evaluated to determine if they should also be considered part of the
object. If so, they are added to the region and the process continues as long
as new pixels are added to the region.  Region growing algorithms vary
depending on the criteria used to decide whether a pixel should be included
in the region or not, the type connectivity used to determine neighbors, and
the strategy used to visit neighboring pixels.

Several implementations of region growing are available in ITK.  This section
describes some of the most commonly used.

\subsection{Connected Threshold}

A simple criterion for including pixels in a growing region is to evaluate
intensity value inside a specific interval.

\label{sec:ConnectedThreshold}
\input{ConnectedThresholdImageFilter.tex}

\subsection{Otsu Segmentation}
Another criterion for classifying pixels is to minimize the error of misclassification.
The goal is to find a threshold that classifies the image into two clusters such that
we minimize the area under the histogram for one cluster that lies on the other cluster's
side of the threshold. This is equivalent to minimizing the within class variance
or equivalently maximizing the between class variance.

\label{sec:OtsuThreshold}
\input{OtsuThresholdImageFilter.tex}

\label{sec:OtsuMultipleThreshold}
\input{OtsuMultipleThresholdImageFilter.tex}

\subsection{Neighborhood Connected}
\label{sec:NeighborhoodConnectedImageFilter}
\input{NeighborhoodConnectedImageFilter.tex}


\subsection{Confidence Connected}
\label{sec:ConfidenceConnected}
\input{ConfidenceConnected.tex}
\input{DesignAndFunctionality/ConfidenceConnectedOnBrainWeb.tex}


\subsection{Isolated Connected}
\label{sec:IsolatedConnected}
\input{IsolatedConnectedImageFilter.tex}


\subsection{Confidence Connected in Vector Images}
\label{sec:VectorConfidenceConnected}
\input{VectorConfidenceConnected.tex}


\section{Segmentation Based on Watersheds}
\label{sec:WatershedSegmentation}
\input{DesignAndFunctionality/WatershedSegmentation.tex}


% the clearpage command helps to avoid orphans in the title of the next
% section.
\clearpage

\section{Level Set Segmentation}
\label{sec:LevelSetsSegmentation}
\input{DesignAndFunctionality/LevelSetsSegmentation.tex}


%HACK: TODO Review what happened to these.
%REMOVED PATENTED: \section{Hybrid Methods}
%REMOVED PATENTED: \label{sec:HybridSegmentationMethods}
%REMOVED PATENTED:
%REMOVED PATENTED: \input{HybridSegmentationMethods.tex}


\section{Feature Extraction}
\label{sec:FeatureExtractionMethods}
\input{DesignAndFunctionality/FeatureExtractionMethods.tex}
